Unbalanced Mallows Models for Optimizing Expensive Black-Box Permutation Problems

### 2021-02-03 paper/main.tex
update heatmaps
updates

The first part of the experimental analysis concerns the {analyses} of the learning rate, which corresponds to parameter $\rho$ of uBorda. As stated in Section~\ref{sec:sampling}, $\rho$ is automatically set at each iteration in such a way that the $100r_1\%$ of the samples have the $100r_2\%$ of the weight. In this regard, we tried a few different values of parameters $r_1$ and $r_2$ in all the instances. Only values of $r_1 < r_2$ make sense and $r_1 = r_2 = 0.5$ would produce equal weights in the learning step. Thus, we tried values $r_1 = \{0.1, 0.2, \dotsc, 0.5\}$ and $r_2 = \{0.6,\dotsc, 0.9, 0.99\}$. Figure~\ref{fig:heatmaps} shows the best solution found by each independent run (averaged over 10 runs) for instances: \texttt{N-t1d100.01} and \texttt{N-t1d150.01} for the LOP, \texttt{reC05} for the PFSP, \texttt{kra32} for the QAP. We terminated runs that took longer than 6 hours and those combinations are shown as white.  Results for all the datasets can be found in the supplemental material~(\supplement).

\begin{figure}[tb]
\centering%
  \begin{minipage}{0.49\linewidth}\centering
    \texttt{N-t59d11xx} (LOP)\\[-0.2ex]
    \includegraphics[width=\textwidth]{../img/heatmap_lop_IO_N-t59d11xx}
  \end{minipage}
  \begin{minipage}{0.49\linewidth}\centering
    \texttt{SGB_N-sgb75_02} (LOP)\\[-0.2ex]
 \includegraphics[width=\textwidth]{../img/heatmap_lop_SGB_N-sgb75_02}
\end{minipage}\\\vskip 1em
\begin{minipage}{0.49\linewidth}
  \centering
  \texttt{reC13} (PFSP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/heatmap_pfsp_rec13_txt}
\end{minipage}
\begin{minipage}{0.49\linewidth}
  \centering
  \texttt{reC19} (PFSP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/heatmap_pfsp_rec19_txt}
\end{minipage}
% \begin{minipage}{0.49\linewidth}
%   \centering
%   \texttt{kra32} (QAP)\\[-0.5ex]
% \includegraphics[width=\textwidth]{../img/heatmap_kra32_dat}
% \end{minipage}
\caption{Mean fitness (over 10 runs) of different configurations of $r_1$ and $r_2$ (darker is better).}\label{fig:heatmaps}}
\end{figure}

@@

\begin{table}[tb]

@@

 \end{table}

\subsection{Rankings and orderings}
%\ekhine{}{rellenar}
In Section~\ref{sec:preliminaries} we highlighted the great difference of rankings and orderings, usually ommitted in the literature. In this lines, we show the consequences of this mistakes. We remind the readear that statistical models (Borda or the Kendall distance) are defined for rankings while standard literature in optimization considers orderings. As a summary of Algorithm~\ref{alg:umm}, we model rankings in our UMM model, generate a ranking $\sigma$, then invert this ranking to obtain an ordering in line~\ref{line:inversion} and evaluate the ordering $\sigma^{-1}$ in the objective function. A natural question is what happens if we did not invert the ranking.

>>>>

The first part of the experimental analysis concerns the {analyses} of the learning rate, which corresponds to parameter $\rho$ of uBorda. As stated in Section~\ref{sec:sampling}, $\rho$ is automatically set at each iteration in such a way that the $100r_1\%$ of the samples have the $100r_2\%$ of the weight. In this regard, we tried a few different values of parameters $r_1$ and $r_2$ in all the instances. Only values of $r_1 < r_2$ make sense and $r_1 = r_2 = 0.5$ would produce equal weights in the learning step. Thus, we tried values $r_1 = \{0.1, 0.2, \dotsc, 0.5\}$ and $r_2 = \{0.6,\dotsc, 0.9, 0.99\}$. Figure~\ref{fig:heatmaps} shows the best solution found by each independent run (averaged over 10 runs) for instances: \texttt{N-t1d100.01} and \texttt{N-t1d150.01} for the LOP, \texttt{reC05} for the PFSP, \texttt{kra32} for the QAP. We terminated runs that took longer than 6 hours and those combinations are shown as white.  Results for all the datasets can be found in the supplemental material~(\supplement).\MANUEL{TODO: Update (mine)}

\begin{figure}[tb]
\centering%
\begin{minipage}{0.49\linewidth}
  \centering
    \texttt{N-t59d11xx} (LOP)\\[-0.2ex]
    \includegraphics[width=\textwidth]{../img/heatmap_lop_IO_N-t59d11xx}
  \end{minipage}
  \begin{minipage}{0.49\linewidth}
    \centering
    \texttt{N-sgb75.02} (LOP)\\[-0.2ex]
 \includegraphics[width=\textwidth]{../img/heatmap_lop_SGB_N-sgb75_02}
\end{minipage}\\\vskip 1em
\begin{minipage}{0.49\linewidth}
  \centering
  \texttt{reC13} (PFSP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/heatmap_pfsp_rec13_txt}
\end{minipage}
\begin{minipage}{0.49\linewidth}
  \centering
  \texttt{reC19} (PFSP)\\[-0.5ex]
  \includegraphics[width=\textwidth]{../img/heatmap_pfsp_rec19_txt}
\end{minipage}
% \begin{minipage}{0.49\linewidth}
%   \centering
%   \texttt{kra32} (QAP)\\[-0.5ex]
% \includegraphics[width=\textwidth]{../img/heatmap_kra32_dat}
% \end{minipage}
\caption{Mean fitness (over 10 runs) of different configurations of $r_1$ and $r_2$ (darker is better).\label{fig:heatmaps}}
\end{figure}

@@

\begin{table*}[tb]

@@

 \end{table*}

\subsection{Rankings and orderings}
In Section~\ref{sec:preliminaries} we highlighted the great difference of rankings and orderings, usually ommitted in the literature. In this lines, we show the consequences of this mistakes. We remind the reader that statistical models (Borda or the Kendall distance) are defined for rankings while standard literature in optimization considers orderings. As a summary of Algorithm~\ref{alg:umm}, we model rankings in our UMM model, generate a ranking $\sigma$, then invert this ranking to obtain an ordering in line~\ref{line:inversion} and evaluate the ordering $\sigma^{-1}$ in the objective function. A natural question is what happens if we did not invert the ranking. 
%\ekhine{}{rellenar}
